Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 interviews.tex
names

%% Raif's usage is typical:
%% \begin{quote}
%% I've been sharing my notebooks in order for other people to see what
%% I've done. It's very convenient for that purpose. [...] We haven't
%% been editing them together, but we've been just sharing, and I've been looking
%% at things that other people do. So there's no collaborative aspect of it, like
%% developing the code, but there's sharing aspects for other people to see and for
%% me to see other people's notebooks.
%% \end{quote}

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is sending links through email. Clint says, ``If my supervisor
wants to see what I've done or QA it, I can just send her a link.''

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Clint says, ``I think the best part about it is how
easily you can share code. [...] You can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Raif
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Taylor develops packages for analysts, and uses RCloud ``more for sharing code with
other people, and for doing tutorials for iotools or hmr'', his packages.
``I want people to see how the package works, so I clearly want them
to see the code [...] I kind of write it just like I would write GitHub
Markdown, where you have little code snippets and text, but RCloud lets me
actually run the snippets'' and display the results. He also uses RCloud for
explaining problems with data sources.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Shelly likes ``the concept of being
able to create notebooks and share them.'' A ``wiki is not the best way for
communicating results - it's kind of like writing a blog post with very limited
functionality [...] I have to save every picture and post it as an image'' and
``I can't share all of the code because it would just get crowded and wouldn't
look right on a wiki.'' Bart says, ``if I could make a folder on RCloud and have
Python notebooks and also Pig notebooks there, and execute them from RCloud,
that would be much better than my current thing, because that would free me from
manual documentation and version control and also telling people where my code
was. It would be just, hey, go look on RCloud, here's the stuff.''

%% Corey: ``being able to share notebooks or codes, or collaborate on something,
%% easily being able to share that information. That's an obvious advantage and
%% nice to have.''

%% Rick: ``I like being able to look at other people's notebooks.''

\subsection{Forking}
The ability to start work where someone else let off, by forking, has proved to
be a popular feature. In fact, almost twice as many (131) users have forked
someone else's notebook as have starred one (75).  Clint says, ``it's one of the
handiest functions, because instead of having to find it, copy, paste it, you
just hit Fork, you rename it, and it's done. It's pretty amazing.''

Although we always intended forking to be used to improve others' code, we we
initially didn't support forking one's own notebooks, and this has proved very
useful. Rick says ``I fork my own notebooks because I'm going off and doing some
other analogous project, so I've got interesting content that I've already done
in a previous analysis that I want to start from and then tweak to match a new
set of data.''

Forking also provides a way for others to troubleshoot when something goes
wrong.  When Horace works with the users of his notebooks, ``I'll teach people
to intercept the result in the middle,'' to ``insert print statements here and
there and check values.''

@@

%% Shelly: ``gives the audience a way to try their own analysis, like maybe Shelly
%% should have used this parameter and they can just change the parameters''

An extremely common but more problematic use of forking is to change
parameters. Raif: ``I've been forking other people's notebooks [because] I
want to run it on a different part of data, or I want to change some parts,
I don't want to see this column, I want a different column, things like
that.''  Taylor complains that a notebook might say ```This is a report of
the volume of all of our feeds for this month', and someone would want to
look at it for the next month or the previous month, so they'd fork it to
change the month.''

Currently parameters can be added to the URL, but adding user interface
elements to do the same thing requires expertise. The tool-builders see a
need for facilities to make it easier. Jyothi calls it ``web-enabling'' the
notebook: users ``can always fork the notebooks and make changes, but I
feel that if the owner of the notebook web-enables it, [...] it's easier to
run. Instead of forking it, if they can set options, it's probably more
efficient, and also they can [still] fork it if they want to.''

%% Clint: ``For the things that I'll do multiple times, I'll try to parameterize
%% it, and I'm working now on putting some packages together that will operate in
%% some fashion outside of RCloud, straight from the command line''

\subsection{Automatic source control}
%% Rick: ``when I've finished something, there's a nice clean record of it.''
%% (I think this may be more about the publish/markdown than about source control.)

Automatic source control is also a popular feature. Clint says, ``instead of
looking back and saying I've got a billion files here in this subdirectory and I
hope I've got them backed up, if they're on RCloud I know they are.''

@@

Bart points out that the automatic versioning works well for dealing with the
minutiae of web development:
\begin{quote}
I like the fact that it has a built-in editor, so
if you need to fix a typo in a link, or an extra line break or any of this other
nonsense, you can just switch to the edit view, pull up your asset, type
something, it's automatically saved, committed, everything. You don't have to go
back to your source code, change it, commit it to the repo, pull the repo to
your distribution version. 
%% None of that's necessary. A lot of web development is
%% fixing stupid little bugs, and you can do that instantly without a lot of
%% overhead, which is nice.
\end{quote}

On the other hand, saving every change leads to many fine-grained
versions. Rick says that for this reason, the history feature is not all that
helpful: ``I don't need something that keeps track of every mistake I've made or
every direction I've tried.''
%% Tagging these versions with names wouldn't
%% necessarily help because ``namespaces are already crowded, so trying to remember
%% the names of notebooks is hard, much less the names of states within each
%% notebook.''

\subsection{Discovering others' work}
Many users find the search function useful. Raif says ``the fact that we have
all the notebooks there, searchable, saves me from replicating what other people
have done.''

But other users prefer to browse just the notebooks of the experts they
know. Rick says, ``usually I know somebody's notebooks that I want to search
through [...] because I know that the kind of thing I'm looking for is something
more obscure than I'm likely to find in some random person's.''

@@

\subsection{Integrated analysis}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull
the data, do the analytics on it, and then feed the viewer the data. So
[...] anyone from the stats group [...] can always insert something in
between. [...] Once you get the data into RCloud, then you have a
dataframe to work with, and then they know how to produce another
dataframe.
\end{quote}

Integrating R also helps Jyothi's application: ``Having an open session
where I can run R commands or R functions without having to invoke an API
or send out a request and then wait for the request to come back [...] is
extremely helpful.''

%% Jyothi: ``another thing I really like about RCloud is your publish feature, so
%% that non-RCloud users, the anonymous login, non-RCloud users can access the
%% notebooks without having to login to RCloud''

@@

This is partly out of familiarity. Taylor asks, ``why would I want to use
RCloud over my current setup? If it's just me, I like my text editor and
terminal. There's nothing that I want that those two don't give me.''

%% Much of this is because of familiarity with other tools. Horace: ``for other
%% people who use RCloud vs RStudio, they will always think that they want to
%% prototype on RStudio. Maybe because it's more familiar for them. So, they work
%% on RStudio, on a small data set, and they fix up everything, make sure
%% everything is right, and then they place it on RCloud for sharing.''

@@

%% But there are interface reasons as well. Rick describes a common workflow:
%% \begin{quote}
%% I have the discipline of having a file that describes what I'm doing, with the
%% commands that I'm using, so that I can go back and recreate it or pass it on to
%% someone else. So it's a little bit like having an RCloud notebook. It's not
%% necessarily executable but most of the commands that I've typed are in there.
%% \end{quote}

The web interface of RCloud doesn't work for Rick: ``It's nice to have
it saved, but there's this trade-off between it making it easier for me to
present something or to save something, and my ease of typing and
correcting and things in a plain old editor window.''

Kenny also works with text file and command line, rearranging a file so the
good code is at the top of the file. When he's done, all the the right is
code at the top. RCloud does not readily support this workflow, but Kenny
often shares the work by pasting it into an RCloud notebook when he's done.

%% Running anything on the Web can also be a source of frustration.  Poor network
%% connections can cause web sockets to die, and Corey complains ``What frustrates me is
%% that if I do leave this window for what feels like 5 or 7 minutes, it
%% disconnects and I have to reconnect and run the whole notebook again.'' Raif
%% whines that ``being web-based, it's obviously slower than something you would be
%% running on your machine.'' For Byers, even the little lag of sending a command
%% to a server and receiving the result is intolerable

Working in a shared environment also entails compromises about what you can
install. Byers says installing alternative or nonstandard packages is
intrinsic to exploratory data analysis.

@@

The interface takes some learning. Clint says at first,
\begin{quote}
%% It took me a couple of weeks of
%% looking at it to become comfortable because 
I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing.
\end{quote}

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick explains:
\begin{quote}
If I was typing into one of the cells near that top, I had
to think ``I'm editing this cell and then I have to execute it,'' and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it.
\end{quote}

%% For some users, the cells getting saved feels incompatible with exploratory data
%% analysis. Rick notes that Markdown is a form of literate programming, and
%% ``being literate about anything typically takes a lot of rewriting and going
%% back over things. [...] I just want to explore a few things and then I'll know
%% what I want to write.''

Much of the time, commands typed at the R command line just serve a transitory
purpose, so having RCloud save commands into cells can be annoying. Kenny notes, ``It saves everything I do like everything is gold, but most
of it is junk not meant to be saved.'' Rick says ``I just want to type in a
couple of quick commands and get some results that are going to tell
me what to do next, and they're not necessarily archival in any sense.''

The material that is not appropriate to save includes ``expressions that allow
me to check that I'm the right track'' (Rick), ``checking out what your data is,
or you make a plot of the data. Things that should not really become part of a
notebook, but things that help you understand your data better'' (Raif).

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces [...] on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell.''

Cells can also take up substantial vertical space, requiring scrolling. Rick
complains that ``every time I type more stuff, the notebook gets longer and
longer and it's harder to deal with.'' Ivan says that cells' controls and blank
space take extra screen space compared to a straight command line, and he and
KC both complained that when there are long results or plots, it causes the
code to scroll off the screen. Raif explains, ``Imagine I run a cell whose
output is huge. [...] To find the next cell, I have to scroll down and find
where that cell starts. So I'm losing the continuity of my code.''

Some users would like the option to keep results separate from code.
Ivan thinks that RStudio's layout is more helpful because charts
are kept in another pane and stay in one place while doing analysis. Raif says
``cells are really useful'' but ``you want to see your output in a different
window [or] on a separate part of the window''.

The cell structure also can be problematic if some cells take a long time to
run. In Kenny's work, there is often a ``long tail''.  The first cell may take
seconds to run, the next cell minutes, and the last cell 5 hours.  In this
situation, the Run Whole Notebook button is ``dangerous''. When converting
his work to notebooks, Kenny ends up with a lot of comments that say ``this cell
takes a long time to run.''

To avert this danger, some users write code within the cells to
cache results. Kenny ends up ``littering the notebook with little switches that
comment out the slow parts'' and either ``load the object instead'' or ``save
the object to disk''. One of the authors of this paper manually writes cells
that check if a result file exists, and perform the calculation only if it's not
there.

%% Rick on auditing (possible solution): ``you didn't see your mistakes, they were
%% in there but you never really saw them because a mistake never led to an answer
%% at the end'' ``the real result of this analysis was these three plots, so go
%% back and figure out everything that I did that was involved in creating those
%% three plots. So that I could start from the raw data and create those three
%% plots.''

%% RCloud combines the functions of scratchpad with notebook assets. KC thinks
%% that having a scratch pad that doesn't persist between different notebooks is
%% ``useless''.

%% The one-window design also makes some users feel that the side panes are taking
%% up too much space. Ivan: having ``long term'' code (which would be in assets)
%% above the command prompt, like in RStudio, means that both can be sized very
%% wide to accomodate long lines.

%% Kenny: Problem with screen real estate: side panes take up too much
%% space. Doesn't need the notebook tree pane.

%% Corey: In the previous version, it switched between code and results. Now you
%% see both - ``and there are times when that's good, but there's certainly times
%% when it'd be nice to - you know, you have 10 or 15 cells and you just want to
%% get back and there's all this output and you're scrolling all over the place so,
%% if there were a way to control both, that'd be great. There are times when I do
%% want to see the output, but there are times when I've seen enough of the same
%% thing.''

%% Jyothi: ``It's good that I can see the results but if I'm working and I need
%% more room and I want to hide the results, I can't.'' Could we have ``an option
%% to close the results window?''

@@

For this reason, Rick doesn't find the notebook tree very useful:
\begin{quote}
I don't necessarily need to see everybody's notebook that uses RCloud [...] Every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.
\end{quote}

Although RCloud promises an environment where notebooks stay working, our
users have not all learned the habits that make this a reality. As Byers
puts it, there is still a big problem with ``bitrot'' -- notebooks often
stop working because the user changed the structure of their data, or
changed a filename or a database.  He says we need ``organizational
protocols'' to catch up with the technology. Even if one forks someone
else's notebook and corrects it, the original notebook still exists with
the error.

@@

Taylor, who writes packages and example notebooks for them, seems to be
haunted the most by dead notebooks. Taylor and Horace work together on a notebook:
\begin{quote}
There's no way for both of us to have ownership of a notebook, so the only
way is to fork it back and forth, and so we have dozens of old copies. [We
end up] deleting all the old ones, but people still have links to them,
because they don't actually disappear.
\end{quote}

The problem of dead notebooks is compounded by changing
packages and their example notebooks at the same time. Taylor will ``go back and
change it and update the page, but then what happens is people have in the
meantime forked it. [...]  I'll have to change a package as well, so their
old fork stops working, and then they complain.''

Taylor tried to keep his notebook tree tidy, but this didn't help, because
people either kept old links in their email, or forked a notebook which is now
obsolete. By design, notebooks that are deleted in the user interface are not
purged from the repository. ``Now I'm really gun-shy about sharing notebooks
because it's like, do I want to support this forever?''

%% Bart: ``One UI notebook and a bunch of other notebooks that do various tasks and
%% are called with API calls. The problem is, if you clone all the notebooks, then
%% you have to go update every single notebook, because all the IDs are different
%% now. There's no sort of relative path type stuff that you can do. You can do it
%% with static assets, but you can't do it with Ajax calls''

>>>>

%% Wendy's usage is typical:
%% \begin{quote}
%% I've been sharing my notebooks in order for other people to see what
%% I've done. It's very convenient for that purpose. [...] We haven't
%% been editing them together, but we've been just sharing, and I've been looking
%% at things that other people do. So there's no collaborative aspect of it, like
%% developing the code, but there's sharing aspects for other people to see and for
%% me to see other people's notebooks.
%% \end{quote}

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is sending links through email. Lilo says, ``If my supervisor
wants to see what I've done or QA it, I can just send her a link.''

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Lilo says, ``I think the best part about it is how
easily you can share code. [...] You can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Wendy
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Evelyn develops packages for analysts, and uses RCloud ``more for sharing code with
other people, and for doing tutorials for iotools or hmr'', his packages.
``I want people to see how the package works, so I clearly want them
to see the code [...] I kind of write it just like I would write GitHub
Markdown, where you have little code snippets and text, but RCloud lets me
actually run the snippets'' and display the results. He also uses RCloud for
explaining problems with data sources.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Leith likes ``the concept of being
able to create notebooks and share them.'' A ``wiki is not the best way for
communicating results - it's kind of like writing a blog post with very limited
functionality [...] I have to save every picture and post it as an image'' and
``I can't share all of the code because it would just get crowded and wouldn't
look right on a wiki.'' Iris says, ``if I could make a folder on RCloud and have
Python notebooks and also Pig notebooks there, and execute them from RCloud,
that would be much better than my current thing, because that would free me from
manual documentation and version control and also telling people where my code
was. It would be just, hey, go look on RCloud, here's the stuff.''

%% Kaylee: ``being able to share notebooks or codes, or collaborate on something,
%% easily being able to share that information. That's an obvious advantage and
%% nice to have.''

%% Kenyon: ``I like being able to look at other people's notebooks.''

\subsection{Forking}
The ability to start work where someone else let off, by forking, has proved to
be a popular feature. In fact, almost twice as many (131) users have forked
someone else's notebook as have starred one (75).  Lilo says, ``it's one of the
handiest functions, because instead of having to find it, copy, paste it, you
just hit Fork, you rename it, and it's done. It's pretty amazing.''

Although we always intended forking to be used to improve others' code, we
initially didn't support forking one's own notebooks, and this has proved very
useful. Kenyon says ``I fork my own notebooks because I'm going off and doing some
other analogous project, so I've got interesting content that I've already done
in a previous analysis that I want to start from and then tweak to match a new
set of data.''

Forking also provides a way for others to troubleshoot when something goes
wrong.  When Hugh works with the users of his notebooks, ``I'll teach people
to intercept the result in the middle,'' to ``insert print statements here and
there and check values.''

@@

%% Leith: ``gives the audience a way to try their own analysis, like maybe Leith
%% should have used this parameter and they can just change the parameters''

An extremely common but more problematic use of forking is to change
parameters. Wendy says, ``I've been forking other people's notebooks [because] I
want to run it on a different part of data, or I want to change some parts,
I don't want to see this column, I want a different column, things like
that.''  Evelyn complains that a notebook might say ```This is a report of
the volume of all of our feeds for this month', and someone would want to
look at it for the next month or the previous month, so they'd fork it to
change the month.''

Currently parameters can be added to the URL, but adding user interface
elements to do the same thing requires expertise. The tool-builders see a
need for facilities to make it easier. Allison calls it ``web-enabling'' the
notebook: users ``can always fork the notebooks and make changes, but I
feel that if the owner of the notebook web-enables it, [...] it's easier to
run. Instead of forking it, if they can set options, it's probably more
efficient, and also they can [still] fork it if they want to.''

%% Lilo: ``For the things that I'll do multiple times, I'll try to parameterize
%% it, and I'm working now on putting some packages together that will operate in
%% some fashion outside of RCloud, straight from the command line''

\subsection{Automatic source control}
%% Kenyon: ``when I've finished something, there's a nice clean record of it.''
%% (I think this may be more about the publish/markdown than about source control.)

Automatic source control is also a popular feature. Lilo says, ``instead of
looking back and saying I've got a billion files here in this subdirectory and I
hope I've got them backed up, if they're on RCloud I know they are.''

@@

Iris points out that the automatic versioning works well for dealing with the
minutiae of web development:
\begin{quote}
I like the fact that it has a built-in editor, so
if you need to fix a typo in a link, or an extra line break or any of this other
nonsense, you can just switch to the edit view, pull up your asset, type
something, it's automatically saved, committed, everything. You don't have to go
back to your source code, change it, commit it to the repo, pull the repo to
your distribution version. 
%% None of that's necessary. A lot of web development is
%% fixing stupid little bugs, and you can do that instantly without a lot of
%% overhead, which is nice.
\end{quote}

On the other hand, saving every change leads to many fine-grained
versions. Kenyon says that for this reason, the history feature is not all that
helpful: ``I don't need something that keeps track of every mistake I've made or
every direction I've tried.''
%% Tagging these versions with names wouldn't
%% necessarily help because ``namespaces are already crowded, so trying to remember
%% the names of notebooks is hard, much less the names of states within each
%% notebook.''

\subsection{Discovering others' work}
Many users find the search function useful. Wendy says ``the fact that we have
all the notebooks there, searchable, saves me from replicating what other people
have done.''

But other users prefer to browse just the notebooks of the experts they
know. Kenyon says, ``usually I know somebody's notebooks that I want to search
through [...] because I know that the kind of thing I'm looking for is something
more obscure than I'm likely to find in some random person's.''

@@

\subsection{Integrated analysis}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Hugh's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull
the data, do the analytics on it, and then feed the viewer the data. So
[...] anyone from the stats group [...] can always insert something in
between. [...] Once you get the data into RCloud, then you have a
dataframe to work with, and then they know how to produce another
dataframe.
\end{quote}

Integrating R also helps Allison's application: ``Having an open session
where I can run R commands or R functions without having to invoke an API
or send out a request and then wait for the request to come back [...] is
extremely helpful.''

%% Allison: ``another thing I really like about RCloud is your publish feature, so
%% that non-RCloud users, the anonymous login, non-RCloud users can access the
%% notebooks without having to login to RCloud''

@@

This is partly out of familiarity. Evelyn asks, ``why would I want to use
RCloud over my current setup? If it's just me, I like my text editor and
terminal. There's nothing that I want that those two don't give me.''

%% Much of this is because of familiarity with other tools. Hugh: ``for other
%% people who use RCloud vs RStudio, they will always think that they want to
%% prototype on RStudio. Maybe because it's more familiar for them. So, they work
%% on RStudio, on a small data set, and they fix up everything, make sure
%% everything is right, and then they place it on RCloud for sharing.''

@@

%% But there are interface reasons as well. Kenyon describes a common workflow:
%% \begin{quote}
%% I have the discipline of having a file that describes what I'm doing, with the
%% commands that I'm using, so that I can go back and recreate it or pass it on to
%% someone else. So it's a little bit like having an RCloud notebook. It's not
%% necessarily executable but most of the commands that I've typed are in there.
%% \end{quote}

The web interface of RCloud doesn't work for Kenyon: ``It's nice to have
it saved, but there's this trade-off between it making it easier for me to
present something or to save something, and my ease of typing and
correcting and things in a plain old editor window.''

Coby also works with text file and command line, rearranging a file so the
good code is at the top of the file. When he's done, all the the right is
code at the top. RCloud does not readily support this workflow, but Coby
often shares the work by pasting it into an RCloud notebook when he's done.

%% Running anything on the Web can also be a source of frustration.  Poor network
%% connections can cause web sockets to die, and Kaylee complains ``What frustrates me is
%% that if I do leave this window for what feels like 5 or 7 minutes, it
%% disconnects and I have to reconnect and run the whole notebook again.'' Wendy
%% whines that ``being web-based, it's obviously slower than something you would be
%% running on your machine.'' For Joy, even the little lag of sending a command
%% to a server and receiving the result is intolerable

Working in a shared environment also entails compromises about what you can
install. Joy says installing alternative or nonstandard packages is
intrinsic to exploratory data analysis.

@@

The interface takes some learning. Lilo says, at first,
\begin{quote}
%% It took me a couple of weeks of
%% looking at it to become comfortable because 
I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing.
\end{quote}

Just the difference between a cell and the command prompt can confuse and slow
down users. Kenyon explains:
\begin{quote}
If I was typing into one of the cells near that top, I had
to think ``I'm editing this cell and then I have to execute it,'' and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it.
\end{quote}

%% For some users, the cells getting saved feels incompatible with exploratory data
%% analysis. Kenyon notes that Markdown is a form of literate programming, and
%% ``being literate about anything typically takes a lot of rewriting and going
%% back over things. [...] I just want to explore a few things and then I'll know
%% what I want to write.''

Much of the time, commands typed at the R command line just serve a transitory
purpose, so having RCloud save commands into cells can be annoying. Coby notes, ``It saves everything I do like everything is gold, but most
of it is junk not meant to be saved.'' Kenyon says ``I just want to type in a
couple of quick commands and get some results that are going to tell
me what to do next, and they're not necessarily archival in any sense.''

The material that is not appropriate to save includes ``expressions that allow
me to check that I'm the right track'' (Kenyon), ``checking out what your data is,
or you make a plot of the data. Things that should not really become part of a
notebook, but things that help you understand your data better'' (Wendy).

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Kenyon says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces [...] on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Evelyn uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell.''

Cells can also take up substantial vertical space, requiring scrolling. Kenyon
complains that ``every time I type more stuff, the notebook gets longer and
longer and it's harder to deal with.'' Gerrard says that cells' controls and blank
space take extra screen space compared to a straight command line, and he and
Tina both complained that when there are long results or plots, it causes the
code to scroll off the screen. Wendy explains, ``Imagine I run a cell whose
output is huge. [...] To find the next cell, I have to scroll down and find
where that cell starts. So I'm losing the continuity of my code.''

Some users would like the option to keep results separate from code.
Gerrard thinks that RStudio's layout is more helpful because charts
are kept in another pane and stay in one place while doing analysis. Wendy says
``cells are really useful'' but ``you want to see your output in a different
window [or] on a separate part of the window''.

The cell structure also can be problematic if some cells take a long time to
run. In Coby's work, there is often a ``long tail''.  The first cell may take
seconds to run, the next cell minutes, and the last cell 5 hours.  In this
situation, the Run Whole Notebook button is ``dangerous''. When converting
his work to notebooks, Coby ends up with a lot of comments that say ``this cell
takes a long time to run.''

To avert this danger, some users write code within the cells to
cache results. Coby ends up ``littering the notebook with little switches that
comment out the slow parts'' and either ``load the object instead'' or ``save
the object to disk''. One of the authors of this paper manually writes cells
that check if a result file exists, and perform the calculation only if it's not
there.

%% Kenyon on auditing (possible solution): ``you didn't see your mistakes, they were
%% in there but you never really saw them because a mistake never led to an answer
%% at the end'' ``the real result of this analysis was these three plots, so go
%% back and figure out everything that I did that was involved in creating those
%% three plots. So that I could start from the raw data and create those three
%% plots.''

%% RCloud combines the functions of scratchpad with notebook assets. Tina thinks
%% that having a scratch pad that doesn't persist between different notebooks is
%% ``useless''.

%% The one-window design also makes some users feel that the side panes are taking
%% up too much space. Gerrard: having ``long term'' code (which would be in assets)
%% above the command prompt, like in RStudio, means that both can be sized very
%% wide to accomodate long lines.

%% Coby: Problem with screen real estate: side panes take up too much
%% space. Doesn't need the notebook tree pane.

%% Kaylee: In the previous version, it switched between code and results. Now you
%% see both - ``and there are times when that's good, but there's certainly times
%% when it'd be nice to - you know, you have 10 or 15 cells and you just want to
%% get back and there's all this output and you're scrolling all over the place so,
%% if there were a way to control both, that'd be great. There are times when I do
%% want to see the output, but there are times when I've seen enough of the same
%% thing.''

%% Allison: ``It's good that I can see the results but if I'm working and I need
%% more room and I want to hide the results, I can't.'' Could we have ``an option
%% to close the results window?''

@@

For this reason, Kenyon doesn't find the notebook tree very useful:
\begin{quote}
I don't necessarily need to see everybody's notebook that uses RCloud [...] Every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.
\end{quote}

Although RCloud promises an environment where notebooks stay working, our
users have not all learned the habits that make this a reality. As Joy
puts it, there is still a big problem with ``bitrot'' -- notebooks often
stop working because the user changed the structure of their data, or
changed a filename or a database.  She says we need ``organizational
protocols'' to catch up with the technology. Even if one forks someone
else's notebook and corrects it, the original notebook still exists with
the error.

@@

Evelyn, who writes packages and example notebooks for them, seems to be
haunted the most by dead notebooks. Evelyn and Hugh work together on a notebook:
\begin{quote}
There's no way for both of us to have ownership of a notebook, so the only
way is to fork it back and forth, and so we have dozens of old copies. [We
end up] deleting all the old ones, but people still have links to them,
because they don't actually disappear.
\end{quote}

The problem of dead notebooks is compounded by changing
packages and their example notebooks at the same time. Evelyn will ``go back and
change it and update the page, but then what happens is people have in the
meantime forked it. [...]  I'll have to change a package as well, so their
old fork stops working, and then they complain.''

Evelyn tried to keep his notebook tree tidy, but this didn't help, because
people either kept old links in their email, or forked a notebook which is now
obsolete. By design, notebooks that are deleted in the user interface are not
purged from the repository. ``Now I'm really gun-shy about sharing notebooks
because it's like, do I want to support this forever?''

%% Iris: ``One UI notebook and a bunch of other notebooks that do various tasks and
%% are called with API calls. The problem is, if you clone all the notebooks, then
%% you have to go update every single notebook, because all the IDs are different
%% now. There's no sort of relative path type stuff that you can do. You can do it
%% with static assets, but you can't do it with Ajax calls''
